---
title: "Understanding Creation of Genetic Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding Creation of Genetic Groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Background
Animals with unknown parents are to be assigned into phantom groups. Phantom groups can be defined based on 

* selection paths (male or female), 
* year of birth of the offspring and
* breed

Possible further grouping criteria can be country or region of origin. 


## Step By Step Analysis
The goal of this analysis is to understand the steps required to define genetic groups. 

### Input
The input to the definition function of genetic groups is a csv-file named `statGenGroupOutFile`. That file comes out of the f90-program `refmtped.f90` which is used to re-format the pedigree. The first few lines of this file is shown below.

```{r, eval=FALSE}
stat_gen_grp_input <- system.file("extdata", "statGenGroupOutFile", package = "qgengroup")
tbl_gen_grp <- readr::read_csv2(file = stat_gen_grp_input, skip = 2)
head(tbl_gen_grp)
```


## Creating Groups With Country
The function `create_GG` is used to create groups that are based on year of birth, breed and country. Inside that function the input is read into a dataframe. Then the function `define_GG` is used to define the genetic groups. 


## Definition of Genetic Groups. 
The definition of genetic groups is done in the function `define_GG`. In a first step different breed labels from the input file can be replaced with a common label. The mapping between the labels is specified in an input file called `psBreedFile`. A similar mapping of labels can also be done with the countries. The groupings of the breeds and the countries is only done if input files for breed- and country mappings are specified.

Unused factor levels are removed after the re-mapping of breeds and countries. 

### Groups on Missing Sires
A new dataframe is defined with only sire selection paths. On that reduced dataframe, the first grouping criterion used are breeds. Breeds with too low group size are separated out and put in a dummy list called `dummy_gg_def_list`. For all breeds with enough missing sires, groups are further subdivided according to country. 


## Definition of Genetic Groups Without Country
The function that creates genetic groups without considering the country of origin of the animals with missing parents is called `define_GG_withoutCountry`. In this function the groupings is done over breeds and year of birth. Groups are formed in the usual way applying some specified lower limits of group sizes.



